---
output: 
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

# log

A logger for R inspired by go's standard library log package.

## Installation

Get it from github.

``` r
# install.packages("remotes")
remotes::install_github("devOpifex/log")
```

## Example

The package comes with a single reference class to create logs.

### Setup

A logger is created from the `Logger` class. Setting `write` to `TRUE` will write every message logged to the `file`. The `prefix` is a string that will precede every message. The `sep` argument is the separator between the `prefix` and the rest of the message to the right, it defaults to a tab (`t`).

```{r}
library(log)

# defaults
log <- Logger$new(
  prefix = "",
  write = FALSE,
  file = "log.log",
  sep = "\t"
)
```

_Different loggers can write to the same file._

One can leave `write` as the default `FALSE` and later use the `dump` method to save it to a file.

### Basic

After the logger has been instantiated one can use the `log` method to log a message to the console (and the file depending on whether that was set).

```{r}
log <- Logger$new(prefix = "INFO")

fnc <- function() {
  log$log("Something")
  Sys.sleep(.7)
  log$log("Something else")
}

fnc()
```

### Flags

When setting up the logger one can customise it with "flags" so it includes a bit more information that might be useful to debug, e.g.: the time and date at which the message was logged. These will be written after the `prefix` and are placed in the order they are used (e.g.: date then time as shown below).

The date and time format can be customised with the `format` argument.

```{r}
errorLog <- Logger$new("ERROR")$
  date()$
  time()

fnc <- function() {
  errorLog$log("Oh no")
  Sys.sleep(.7)
  errorLog$log("Snap!")
}

fnc()
```

You can also customise the look of the prefix with `hook`, pass it a function that will take the prefix and return a modified version of it.

While the package comes with basic flags you can add your own with the `flag` method. This method accepts either a function that will be run every time a message is logged or a string that will simply be included in the message.

```{r}
random_flag <- function(){
  paste0(sample(letters, 3), collapse = "|")
}

l <- Logger$new("ERROR")$
  time()$
  flag("[README.md]")$
  flag(random_flag)

fnc <- function() {
  l$log("Oh no")
  Sys.sleep(.7)
  l$log("Snap!")
}

fnc()
```

### Hook

There is also the possibility to pass a "hook;" a function that will preprocess the prefix and return a modified version of it.

```{r}
# using crayon
errorLog <- errorLog$hook(crayon::red)

# fancier
hook <- function(x) {
  paste(crayon::blue(cli::symbol$info, x))
}

log <- Logger$new("INFO")$hook(hook)
log$log("Fancy this!?")
```

### Dump

Finally you can dump the log to a file with `dump`.

```r
log$dump("stuff.log")
```
